---
title: "Analyze_Cyclistic_Part1"
author: "Jiejun"
date: "2023-12-09"
output: html_document
---

## Analyze Part 1
In this R Markdown, I will do the steps to analyze the data (part 1). Most of this will be based on this Google Doc here -> [link](https://docs.google.com/document/d/1ZE95TYd94HN7DQ4_Mfg4SAMYEXoRmNlNG_uvUlzA_Jo/edit?usp=sharing)

Now we take a look at the data we processed earlier using some functions from packages

### Install packages

```{r install packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(skimr)
```

### Read the data we processed earlier

```{r read data, message=FALSE, warning=FALSE}
all_trips <- read_csv("data/all_trips_processed.csv")
```

### View the data

```{r colnames all_trips}
colnames(all_trips)
```
There are 8 column names as shown above

```{r head all_trips}
head(all_trips)
```

The data seems fine based on the head of the data

```{r skim_without_charts all_trips}
skim_without_charts(all_trips)
```

There seems to still be 1 missing row for the end_station id and name. Now we check the data types

```{r}
str(all_trips)
```

Looks good to me so far

### Add a day, month, year column for the start_date

In order to examine trends over this 1 year period, we can separate the start dates (started_at) into day, month and year. The doc gives us a link that we can follow to learn more -> [link](https://www.statmethods.net/input/dates.html)

```{r}
all_trips$start_date <- as.Date(all_trips$started_at)
```

```{r}
all_trips$start_month <- format(as.Date(all_trips$start_date), "%m")
all_trips$start_day <- format(as.Date(all_trips$start_date), "%d")
all_trips$Start_year <- format(as.Date(all_trips$start_date), "%Y")
all_trips$Start_day_of_week <- format(as.Date(all_trips$start_date), "%A")

```

We should also add a ride length in seconds so that we can analyze that in our trend analysis in the future

* from this point on, all my data suddenly disappeared, but my csv is still made somehow. Following these steps should still give you the same conclusion. I think R may have taken a bit too much memory and it didn't save correctly. Maybe I should have just closed the other tabs

anyway

### Add a ride duration

Click here -> [link](https://stat.ethz.ch/R-manual/R-devel/library/base/html/difftime.html) for more information

```{r}
all_trips$ride_length <- difftime(all_trips$ended_at,all_trips$started_at)
```

Convert the ride lengths to numeric

```{r}
all_trips$ride_length <- as.numeric(all_trips$trip_dur)
```

Remove the bad data 

The dataframe includes a few hundred entries when bikes were taken out of docks and checked for quality by Divvy or ride_length was negative

We will create a new version of the dataframe (v2) since data is being removed

Click here -> [link](https://www.datasciencemadesimple.com/delete-or-drop-rows-in-r-with-conditions-2/) for more information

```{r}
all_trips_v2 <- all_trips[!(all_trips$start_station_name == "HQ QR" | all_trips$ride_length<0),]
```

Now we write all the data into a csv

```{r}
write_csv(all_trips_v2, file = 'C:\\Users\\jieju\\Desktop\\GoogleDA_CaseStudy\\data\\all_trips_transformed.csv')
```

We move to part 2 of the analysis. I'm gonna save and push to github real quick